---
title: "STATS 2DA3 Assignment 5"
author: "Faris Abuain"
date: "2024-11-26"
output: pdf_document
geometry: margin=1.0in
mainfont: Times New Roman
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "https://cran.rstudio.com/"))
```

1. Using the Assign5.csv dataset (.csv file) that can be found on the Avenue to Learn page under “Assignments”, complete the following tasks: (Note: you may need to load multiple libraries to answer the questions.)

• use the default initialization and stopping conditions of the algorithm in question. Set the seed to 1234 whenever a seed is needed to ensure reproducibility of results.


```{r}
ass_5_data <- read.csv("~/Downloads/Assign5.csv")
head(ass_5_data)
str(ass_5_data)
```


a) Make the “Class” variable into a factor.

```{r}
ass_5_data$Class <- as.factor(ass_5_data$Class)
```


b) Remove the “Class” variable and scale the dataset.

```{r}
ass_5_no_class <- ass_5_data[, -ncol(ass_5_data)]

scaled_ass_data <- scale(ass_5_no_class)
head(scaled_ass_data)
```


c) Carry out Model-based clustering using Mclust with the ultimate goal to predict
“Class”, using all remaining variables to produce your model. The model should
consider 1 to 4 components (clusters).


```{r, echo=FALSE}
install.packages("mclust")
```


```{r} 
library(mclust)
set.seed(1234)

mclust_model <- Mclust(scaled_ass_data,1:4)
```


i. According to the BIC, how many components give the best result?

```{r}
best_result_BIC <- mclust_model$G
best_result_BIC
```

Therefore 4 components gives the best result.


ii. Produce a graph showing the BICs for all fitted models as a function of the number
of components.

```{r}
library(ggplot2)

bic_vals <- mclust_model$BIC
num_components <- 1:4
bic_df <- data.frame(Components = num_components, BIC = bic_vals[1:4])


bic_plot <- ggplot(bic_df, aes(x = Components, y = BIC))
bic_plot <- bic_plot + geom_line() 
bic_plot <- bic_plot + geom_point()
bic_plot <-  bic_plot + labs(title = 
                        "BIC Values for Fitted Models with Varying Number of Components", x = "Number of Components", y = "BIC") 
bic_plot <- bic_plot + theme(plot.title = element_text(hjust = 0.5))

bic_plot
```


d) Carry out Model-based clustering using gpcm with the ultimate goal to predict “Class”, using all remaining variables to produce your model. The model should consider 1 to 4 components (clusters).


```{r}
install.packages("mixture")
```



```{r}
library(mixture)
set.seed(1234)

gpcm_model <- gpcm(scaled_ass_data, G = 1:4)  
```


i. Produce a table of results, showing actual class against predicted class.


```{r}
actual_class <- ass_5_data$Class

predicted_class <- gpcm_model$map

results_table <- table(predicted_class, actual_class)
results_table
```


ii. What is the ARI correct to 4 decimal places?

```{r}
library(e1071)
agreement_value <- classAgreement(results_table)
ari_2 <- agreement_value$crand
ari_rounded <- round(ari_2,4)
ari_rounded
```

Therefore the adjusted rand index (ARI) is 0.8645 for our gpcm model. 


2. 


a) Set the seed to (1234) and generate 20 observations from a normal distribution with mean 4 and standard deviation 9. (Show the output)

```{r}
set.seed(1234)
observations <- rnorm(20,4,9)
print(observations)
```


b) Calculate the mean of these 20 numbers (correct to 3 places of decimals).

```{r}
mean_value <- round(mean(observations), 3)
mean_value
```


c) Using the same seed, calculate a boot-strapped estimate of the mean of these 20
numbers (correct to 3 places of decimals), using 50 bootstrapped samples. Use the
sample function as part of your coding to carry out this task.


```{r}
bootstrap_means <- replicate(50, mean(sample(sample(observations, length(observations), replace = TRUE))))

bootstrap_mean_rounded <- round(mean(bootstrap_means), 3)
bootstrap_mean_rounded
```

d) What is the bias of the boot-strapped estimate of the mean?


```{r}
bias <- bootstrap_mean_rounded - mean_value
bias
```

Thus the bias of the boot-strapped estimate is -0.112. 


END OF ASSIGNMENT. 